---
title: "PCA & GLMS for for Community Data Paper"
date: "2023-05-12"
output: html_document
---

This script was run with R 4.2.2

# Libraries

```{r load libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lubridate)
library(stringi)
library(stringr)
library(dplyr)
library(reshape2)
library(tidyr)
library(ggplot2)

library(ggpmisc)
library(ggpubr) 
library(ggthemes)
library(corrplot)


```
#Monthly Surveys
##Combine Data
``````{r Monthly Combine, include=FALSE}
#####Monthly Light Data######

IRM <- read.csv("~/Monthly_Light.csv")
colnames(IRM)
IRM$site.point<-paste(IRM$Site, " ", IRM$IRR.Point, sep="")

#Rename by Site
IRM$site.point<-str_replace(IRM$site.point, "Woolsey 1", "Woolsey")
IRM$site.point<-str_replace(IRM$site.point, "Woolsey 2", "Woolsey")
IRM$site.point<-str_replace(IRM$site.point, "Woolsey 3", "Woolsey")
IRM$site.point<-str_replace(IRM$site.point, "Woolsey 4", "Woolsey")
IRM$site.point<-str_replace(IRM$site.point, "Woolsey 5", "Woolsey")
colnames(IRM)

#calculate relative UV
IRM$relative.uv <- IRM$mean.uv.auc/IRM$mean.auc

#extract month info
IRM$month<- substr(IRM$Date,5,7)
IRM$month <- factor(IRM$month, levels = month.abb)
summary(IRM)

#sort by BGO sunny, shade, & butterfly house sites
IRM$site.simple<-IRM$site.point
IRM$site.simple<-str_replace(IRM$site.simple, "BGO 1", "BGO Sunny")
IRM$site.simple<-str_replace(IRM$site.simple, "BGO 2", "BGO Sunny")
IRM$site.simple<-str_replace(IRM$site.simple, "BGO 3", "Butterfly_House")
IRM$site.simple<-str_replace(IRM$site.simple, "BGO 4", "BGO Shade")
IRM$site.simple<-str_replace(IRM$site.simple, "BGO 5", "BGO Shade")

#Remove Butterfly House Data (not part of study)
IR.Final<-filter(IRM, site.simple != "Butterfly_House")

#match months with abbreviations
IR.Final$abb<-match(IR.Final$month, month.abb)


#######Combine Monthly Count & Irrad######

#create survey ID to match with abundance data
IR.Final$Survey.ID<-paste(IR.Final$Date, IR.Final$Site, sep = " ")

#find mean total irradiance for each survey visit
mo.light <- aggregate(IR.Final$mean.auc, 
                      by = list(IR.Final$Survey.ID), FUN = mean, na.rm=F)
#find mean relative UV for each survey visit
mo.uv <- aggregate(IR.Final$relative.uv, 
                   by = list(IR.Final$Survey.ID), FUN = mean, na.rm=F)

#rename columns
names(mo.light)[1] <- "Survey.ID"
names(mo.light)[2] <- "total_irrad"
names(mo.uv)[1] <- "Survey.ID"
names(mo.uv)[2] <- "rel.uv"

#join average total irradiance & relative uv dataframes
mo.most2<-left_join(mo.light,mo.uv, by=c("Survey.ID"))

#Get abundance & weather data
mo.counts <- read.csv("~/Monthly_Abundance.csv")


#Create Survey ID
mo.counts$Site<-str_replace(mo.counts$Site, "Botanical_Garden", "BGO")
mo.counts$Survey.ID<-paste(mo.counts$Date, mo.counts$Site, sep = " ")
colnames(mo.counts)
#tidy the data
mo.melt = melt(mo.counts, id.vars = c("Survey.ID", "Site", "Date", "Month", "Year", "total_butterflies", "TempF", "Weekly_Precip", "Weekly_Precip_cm"),
               measure.vars = c("nymphalidae", "pieridae", "papilionidae", "hesperiidae", "lycaenidae"))
names(mo.melt)[10] <- "Family"
names(mo.melt)[11] <- "Count"
mo.melt$Family<-str_replace(mo.melt$Family, "nymphalidae", "Nymphalidae")
mo.melt$Family<-str_replace(mo.melt$Family, "pieridae", "Pieridae")
mo.melt$Family<-str_replace(mo.melt$Family, "hesperiidae", "Hesperiidae")
mo.melt$Family<-str_replace(mo.melt$Family, "papilionidae", "Papilionidae")
mo.melt$Family<-str_replace(mo.melt$Family, "lycaenidae", "Lycaenidae")
summary(mo.light)
summary(mo.light$Survey.ID)

#join count, weather & light data
mo<-left_join(mo.melt, mo.most2, by=c("Survey.ID"))

#add numerical months
mo$abb<-match(mo$Month, month.name)
#Convert to Celsius
mo$TempC<-(mo$TempF-32)/ 1.8 
```
##Correlation Matrix
```{r lCorrelation matrix, include=TRUE}

#convert site(factor) into number to plug into correlation matrix

mo$site.num<-mo$Site 
mo$site.num<-str_replace(mo$site.num, "Woolsey", "2")
mo$site.num<-str_replace(mo$site.num, "BGO", "1")
mo$site.num<-as.numeric(mo$site.num)
colnames(mo)

#include Year, Weekly Precipitation (cm) , Total Irradiance, Relative UV, Abbreviated Month (abb), Temperature (C), & Site
mo.matx <- cor(na.omit(mo[ , c(6,10,13,14,15,16,17)]))

#Nice optional visualization
corrplot(mo.matx, method="pie")

mo.matx.rounded <- round(x= mo.matx, digits=3)
View(mo.matx.rounded)
write.csv(mo.matx.rounded, "month_corr_MayTest.csv", row.names=TRUE)
```

## PCA

Runs the Principal Components analysis and shows how much each component explains variance in the data

```{r PCA for Monthly Surveys, include=TRUE}
#Site was left out of PCA (no column 17) for later analysis of effect of site
mo.pca <- princomp(na.omit(mo[ , c(6,10,13,14,15,16)]),
                           scores = TRUE,
                           cor = TRUE
)

#proportion of variance table
summary(mo.pca)
```

Details the loadings for each Principal Component ("abb" represents survey month as a numeric14

```{r Details of PCA for Monthly Surveys, include=TRUE}
loadings.mo <- unclass(mo.pca[["loadings"]])
sweep (loadings.mo,
       MARGIN = 2,
       STATS = colSums(loadings.mo),
       FUN = "/")
```

Biplot & Scree Plots for monthly data PCA

```{r Monthly PCA, include=TRUE}
#visualize Principal Components
biplot(mo.pca)
screeplot(x = mo.pca,
          npcs = length(mo.pca[["sdev"]]),
          type = "lines")
```

Preparing the data for generalized linear models:

```{r Monthly PCA, include=FALSE}
#Bind together PCs of interest
moPC1 <- data.frame(mo.pca[["scores"]])[,1]
moPC2 <- data.frame(mo.pca[["scores"]])[,2]
moPC3<- data.frame(mo.pca[["scores"]])[,3]
moscores1<-cbind(moPC1, moPC2)
moscores<-cbind(moscores1, moPC3)

#Bind abundance data with PC scores
mo1<-na.omit(mo)
mo.PCs <- cbind(mo1, moscores)


```

## GLM Visualizations

```{r Visualizing Generalized Linear Models, include=TRUE}

##Prepare the data to create faceted figure
melt.mo <- melt(mo.PCs, id.vars = c("Site", "Date", "Year", "Month", "Survey.ID","total_butterflies"),
            measure.vars = c("moPC1", "moPC2", "moPC3"))

melt.mo.total <- melt(mo.PCs, id.vars = c("Site", "Date", "Year", "Month", "Survey.ID","total_butterflies"),
                    measure.vars = c("moPC1", "moPC2", "moPC3"))
melt.mo.total$Family <- "All Families"
names(melt.mo.total)[6] <- "Count"

melt.mo.fam <- melt(mo.PCs, id.vars = c("Site", "Date", "Year", "Month", "Survey.ID","Family","Count"),
                measure.vars = c("moPC1", "moPC2", "moPC3"))
mo.facet<-rbind(melt.mo.fam,melt.mo.total)
mo.facet$Family <- factor(mo.facet$Family , levels=c("All Families", "Hesperiidae", "Lycaenidae", "Nymphalidae", "Papilionidae", "Pieridae"))

mo.facet2<-mo.facet %>%
  distinct(.keep_all = TRUE)


##Visualize: Family vs PC linear regressions
# R^2 values appear on graphs
q<-ggplot(data = mo.facet2, aes(x = value, y = Count)) +
  stat_poly_line() +
  stat_poly_eq() +
  geom_point(size=3) +
  theme_classic() +
  annotate(
    geom = 'segment',
    y = -Inf,
    yend = Inf,
    x = Inf,
    xend = Inf
  ) + 
  theme(strip.placement = "outside")
facet(q, facet.by = c("Family","variable"),
      scales = "free",
      panel.labs.background = list(size = 0.1),
      panel.labs.font = list(color = "black"),
      panel.labs.font.x = list( color = "black"))

#Visualize: Family vs Site boxplot

ggplot(mo.facet2,aes(Family, Count, fill=Site)) +
  stat_boxplot(size=2,outlier.size=5) + 
  theme_classic() +  
  theme(legend.title = element_blank(),legend.text=element_text(size=25), 
        axis.text=element_text(size=10, angle = 45, hjust=1, face="bold"),
        axis.text.x=element_text(margin = margin(t=0, b=0)),
        axis.title.y = element_text(size=28,face="bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size=28,face="bold",
                                    margin = margin(t = 20, r = 0, b = 0, l = 0)),
        plot.title = element_text(hjust = 0.5, size=32, face="bold")) +
  scale_fill_manual(values = c('#ffc000', '#0072b2')) +
  labs(x="Family", y= "# of Butterflies")
```

## GLMs

```{r Generalized Linear Models - Monthly Surveys , include=TRUE}
#filter for each family
mo.hes<-filter(mo.PCs,Family=="Hesperiidae")
mo.pap<-filter(mo.PCs,Family=="Papilionidae")
mo.nym<-filter(mo.PCs,Family=="Nymphalidae")
mo.pie<-filter(mo.PCs,Family=="Pieridae")
mo.lyc<-filter(mo.PCs,Family=="Lycaenidae")

mo.Pcsd<-mo.PCs %>%
  distinct(Survey.ID, .keep_all = TRUE)

#GLM for all butterflies  
all.glm <- glm(total_butterflies ~ moPC1+moPC2+moPC3 + Site, data=mo.Pcsd, family=quasipoisson)
summary(all.glm)
anova(all.glm, test = "Chisq")
```

```{r Generalized Linear Models - Monthly Surveys , include=TRUE}
#GLM for Hesperiidae
hes.glm <- glm(Count ~ moPC1+moPC2+moPC3 + Site, data=mo.hes, family=quasipoisson)
summary(hes.glm)
anova(hes.glm, test = "Chisq")
```

```{r Generalized Linear Models - Monthly Surveys , include=TRUE}
#GLM for Lycaenidae
lyc.glm <- glm(Count ~ moPC1+moPC2+moPC3 + Site, data=mo.lyc, family=quasipoisson)
summary(lyc.glm)
anova(lyc.glm, test = "Chisq")
```

```{r Generalized Linear Models - Monthly Surveys , include=TRUE}
#GLM for Nymphalidae 
nym.glm <- glm(Count ~ moPC1+moPC2+moPC3 + Site, data=mo.nym, family=quasipoisson)
summary(nym.glm)
anova(nym.glm, test = "Chisq")
```

```{r Generalized Linear Models - Monthly Surveys , include=TRUE}
#GLM Papilionidae
pap.glm <- glm(Count ~ moPC1+moPC2+moPC3 + Site, data=mo.pap, family=quasipoisson)
summary(pap.glm)
anova(pap.glm, test = "Chisq")
```

```{r Generalized Linear Models - Monthly Surveys , include=TRUE}
#GLM for Pieridae
pie.glm <- glm(Count ~ moPC1+moPC2+moPC3 + Site, data=mo.pie, family=quasipoisson)
summary(pie.glm)
anova(pie.glm, test = "Chisq")
```

#Biweekly Surveys
## Combine Data
```{r Combine Biweekly Data, include=False}
#####Biweekly Irrad Measurements#######
IR1.2 <- read.csv("~/Upload_Biweekly_Light_Upload.csv")

IR1.2$Survey.ID <-paste(IR1.2$Date, " ", IR1.2$Site,  sep="")
IR1.2$Irr.ID <-str_sub(IR1.2$folder.names,-1,-1)

#agg by site point
auc.cp <- aggregate(IR1.2$mean.auc,
                    by = list(IR1.2$Week, IR1.2$Site, IR1.2$Survey.ID ) , FUN = mean, na.rm=FALSE)

IR1.2$relative.uv <- IR1.2$mean.uv.auc/IR1.2$mean.auc
rel.uv.cp <- aggregate(IR1.2$relative.uv,
                       by = list(IR1.2$Week, IR1.2$Site, IR1.2$Survey.ID ) , FUN = mean, na.rm=FALSE)

IR.cp<-left_join(auc.cp,rel.uv.cp, by=c("Group.1","Group.2","Group.3"))
colnames(IR.cp)
IR.cp

names(IR.cp)[1] <- "Week"
names(IR.cp)[2] <- "Site"
names(IR.cp)[3] <- "Survey.ID"
names(IR.cp)[4] <- "auc"
names(IR.cp)[5] <- "rel.uv"


summary(IR.cp)


###########Count & Weather Data######
#bi <- read.csv("~/Research/Community Data Paper/Data/transect_3years.csv")
bi <- read.csv("~/Upload_Biweekly_Count_Weather.csv")

bi$Survey.ID<-paste(bi$Date, bi$Site, sep = " ")

bi.melt = melt(bi, id.vars = c("Survey.ID", "Site", "Date", "Time.of.Day", "Year", "Julian_Week", "Total", "Temp", "This_week.s_total_precip_in", "Transect"),
            measure.vars = c("Nymphalidae", "Pieridae", "Papilionidae", "Hesperiidae", "Lycaenidae"))

######Temp Data######

bi$TempC<-(bi$Temp-32)/ 1.8
bi.temp <- aggregate(bi$TempC, 
  by = list(bi$Survey.ID, bi$Julian_Week) , FUN = mean, na.rm=FALSE)

names(bi.temp)[1] <- "Survey.ID"
names(bi.temp)[2] <- "Week"
names(bi.temp)[3] <- "TempC"

bi.temp$Week.factor<-as.factor(bi.temp$Week)



#####Precip Data######
bi.precip <- aggregate(bi$This_week.s_total_precip_in, 
                     by = list(bi$Year, bi$Julian_Week) , FUN = mean, na.rm=FALSE)


names(bi.precip)[1] <- "Year"
names(bi.precip)[2] <- "Week"
names(bi.precip)[3] <- "Precip"

bi.precip$PrecipCM<-bi.precip$Precip*2.54


#bi.precip$Julian_Week<-as.factor(bi.precip$Julian_Week)
bi.precip$Weekf<-as.factor(bi.precip$Week)

########## Combine Count. Light & Weather######
bi.light <- aggregate(bi.melt$value, 
                     by = list(bi.melt$Year, bi.melt$Julian_Week, bi.melt$Site, bi.melt$Date, bi.melt$variable), FUN = sum, na.rm=FALSE)
bi.light2 <- aggregate(bi.melt$Total, 
                      by = list(bi.melt$Year, bi.melt$Julian_Week, bi.melt$Site, bi.melt$Date, bi.melt$variable), FUN = sum, na.rm=FALSE)

names(bi.light)[1] <- "Year"
names(bi.light)[2] <- "Week"
names(bi.light)[3] <- "Site"
names(bi.light)[4] <- "Date"
names(bi.light)[5] <- "Family"
names(bi.light)[6] <- "Fam.Count"

names(bi.light2)[1] <- "Year"
names(bi.light2)[2] <- "Week"
names(bi.light2)[3] <- "Site"
names(bi.light2)[4] <- "Date"
names(bi.light2)[5] <- "Family"
names(bi.light2)[6] <- "Total.Count"

bi.light$Survey.ID<-paste(bi.light$Date, bi.light$Site, sep = " ")
bi.light2$Survey.ID<-paste(bi.light2$Date, bi.light2$Site, sep = " ")
bi.most<-left_join(bi.light,bi.light2, by=c("Survey.ID","Week","Year", "Site", "Date", "Family"))

bi.temp$Week<-as.numeric(bi.temp$Week)
bi.precip$Week<-as.numeric(bi.precip$Week)
bi.most2<-left_join(bi.most,bi.temp, by=c("Survey.ID","Week"))
bi.most3<-left_join(bi.most2,bi.precip, by=c("Year","Week"))


bi.all<-left_join(bi.most3,IR.cp, by=c("Survey.ID", "Week", "Site"))

summary(bi.all)

```
##Correlation Matrix
```{r Correlation matrix, include=TRUE}

#convert site(factor) into number to plug into correlation matrix

bi.all$site.num<-bi.all$Site
bi.all$site.num<-str_replace(bi.all$site.num, "Woolsey", "1")
bi.all$site.num<-str_replace(bi.all$site.num, "Stump", "2")
bi.all$site.num<-str_replace(bi.all$site.num, "Chesney", "3")
bi.all$site.num<-as.numeric(bi.all$site.num)

#include Year, Weekly Precipitation (cm) , Total Irradiance (auc), Relative UV, Survey Week, Temperature (C), & Site
bi.matx <- cor(na.omit(bi.all[ , c(1,2,9,12,14,15,16)]))

#Nice optional visualization
corrplot(bi.matx, method="pie")

bi.matx.rounded <- round(x= bi.matx, digits=3)
View(bi.matx.rounded)
write.csv(bi.matx.rounded, "week_corr_MayTest.csv", row.names=TRUE)
```
##PCA
```{r PCA for Biweekly Surveys, include=TRUE}
#Site was left out of PCA (no column 16) for later analysis of effect of site
bi.pca <- princomp(na.omit(bi.all[ , c(1,2,9,12,14,15)]),
                   scores = TRUE,
                   cor = TRUE
)
#proportion of variance table
summary(bi.pca)
```

```{r Details of PCA for Biweekly Surveys, include=TRUE}
loadings.bi <- unclass(bi.pca[["loadings"]])
sweep (loadings.bi,
       MARGIN = 2,
       STATS = colSums(loadings.bi),
       FUN = "/")
```
Biplot & Scree Plots for biweekly data PCA

```{r Biweekly PCA, include=TRUE}
#visualize Principal Components
biplot(bi.pca)
screeplot(x = bi.pca,
          npcs = length(bi.pca[["sdev"]]),
          type = "lines")
```

Preparing the data for generalized linear models:

```{r Biweekly PCA, include=FALSE}
#Bind together PCs of interest
biPC1 <- data.frame(bi.pca[["scores"]])[,1]
biPC2 <- data.frame(bi.pca[["scores"]])[,2]
biPC3 <- data.frame(bi.pca[["scores"]])[,3]
biPC4 <- data.frame(bi.pca[["scores"]])[,4]
biscores<-cbind(biPC1, biPC2, biPC3, biPC4)

#Bind abundance data with PC scores
bi1<-na.omit(bi.all)
bi.PCs <- cbind(bi1, biscores)


```
## GLM Visualizations

```{r Visualizing Generalized Linear Models, include=TRUE}

##Prepare the data to create faceted figure
melt.bi.total <- melt(bi.PCs, id.vars = c("Site", "Date", "Year", "Week", "Survey.ID","Total.Count"),
                      measure.vars = c("biPC1", "biPC2", "biPC3", "biPC4"))
melt.bi.total$Family <- "All Families"
names(melt.bi.total)[6] <- "Fam.Count"

melt.bi.fam <- melt(bi.PCs, id.vars = c("Site", "Date", "Year", "Week", "Survey.ID","Family","Fam.Count"),
                    measure.vars = c("biPC1", "biPC2", "biPC3","biPC4"))
bi.facet<-rbind(melt.bi.fam,melt.bi.total)
bi.facet$Family <- factor(bi.facet$Family , levels=c("All Families","Hesperiidae", "Lycaenidae", "Nymphalidae", "Papilionidae", "Pieridae"))

bi.facet2<-bi.facet %>%
  distinct(.keep_all = TRUE)


##Visualize: Family vs PC linear regressions
# R^2 values appear on graphs
q<-ggplot(data = bi.facet, aes(x = value, y = Fam.Count)) +
  stat_poly_line() +
  stat_poly_eq() +
  geom_point(size=3) +
  theme_base()+
  theme(strip.placement = "outside")

facet(q, facet.by = c("Family","variable"),
      scales = "free",
      panel.labs.background = list(size = 0.5),
      panel.labs.font = list(color = "black"),
      panel.labs.font.x = list( color = "black"))

#Visualize: Family vs Site boxplot
ggplot(bi.facet2,aes(x=factor(Family, level=c("All Families", "Hesperiidae", "Lycaenidae", "Nymphalidae", "Papilionidae", "Pieridae")),Fam.Count, fill=Site)) +
  stat_boxplot(size=2,outlier.size=5) + 
  theme_classic() +  
  theme(legend.title = element_blank(),legend.text=element_text(size=25), 
        axis.text=element_text(size=28, face="bold"),
        axis.title.y = element_text(size=28,face="bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size=28,face="bold",
                                    margin = margin(t = 20, r = 0, b = 0, l = 0)),
        plot.title = element_text(hjust = 0.5, size=32, face="bold")) +
  scale_fill_manual(values = c('#d55e00', '#f0e442', '#0072b2')) +
  labs(x="Family", y= "# of Butterflies")
```
## GLMs

```{r Generalized Linear Models - Biweekly Surveys, include=TRUE}
#filter for each family

bi1.hes<-filter(bi.PCs,Family=="Hesperiidae")
bi1.pap<-filter(bi.PCs,Family=="Papilionidae")
bi1.nym<-filter(bi.PCs,Family=="Nymphalidae")
bi1.pie<-filter(bi.PCs,Family=="Pieridae")
bi1.lyc<-filter(bi.PCs,Family=="Lycaenidae")

bi.Pcsd<-bi.PCs %>%
  distinct(Survey.ID, .keep_all = TRUE)

#GLM for all butterflies  
bi.tot <- glm(Total.Count ~ biPC1 + biPC2 + biPC3 + biPC4 + Site, data=bi.Pcsd, family=quasipoisson)
summary(bi.tot)
anova(bi.tot, test = "Chisq")
```

```{r Generalized Linear Models - Biweekly Surveys, include=TRUE}
#GLM for Hesperiidae
bi.hes.glm<-glm(Fam.Count~biPC1 + biPC2 + biPC3 + biPC4 + Site, data=bi1.hes, family=quasipoisson)
summary(bi.hes.glm)
anova(bi.hes.glm, test = "Chisq")
```

```{r Generalized Linear Models - Biweekly Surveys, include=TRUE}
#GLM for Lycaenidae
bi.lyc.glm<-glm(Fam.Count~biPC1 + biPC2 + biPC3 + biPC4 + Site, data=bi1.lyc, family=quasipoisson)
summary(bi.lyc.glm)
anova(bi.lyc.glm, test = "Chisq")
```

```{r Generalized Linear Models - Biweekly Surveys, include=TRUE}
#GLM for Nymphalidae 
bi.nym.glm<-glm(Fam.Count~biPC1 + biPC2 + biPC3 + biPC4 + Site, data=bi1.nym, family=quasipoisson)
summary(bi.nym.glm)
anova(bi.nym.glm, test = "Chisq")
```

```{r Generalized Linear Models - Biweekly Surveys, include=TRUE}
#GLM Papilionidae
bi.pap.glm<-glm(Fam.Count~biPC1 + biPC2 + biPC3 + biPC4 + Site, data=bi1.pap, family=quasipoisson)
summary(bi.pap.glm)
anova(bi.pap.glm, test = "Chisq")
```

```{r Generalized Linear Models - Biweekly Surveys , include=TRUE}
#GLM for Pieridae
bi.pie.glm<-glm(Fam.Count~biPC1 + biPC2 + biPC3 + biPC4 + Site, data=bi1.pie, family=quasipoisson)
summary(bi.pie.glm)
anova(bi.pie.glm, test = "Chisq")
```
#Behavior
```{r Generalized Linear Models - Biweekly Surveys Behavior , include=TRUE}


bi.beh <- read.csv("~/processed_beh_transects_2023Jan4.csv")
bi.beh$Survey.ID<-paste(bi.beh$Date, bi.beh$Site, sep = " ")
bi.beh$Year<-as.numeric(bi.beh$Year)

bi.combo<-left_join(bi.beh,bi.PCs, by=c("Survey.ID","Date","Year","Site","Family","Week"))
summary(bi.combo$Behavior)
bi.combo$Behavior<-as.factor(bi.combo$Behavior)




#File for JMP
melt.beh <- melt(bi.combo, id.vars = c("Site", "Date", "Year", "Week", "Survey.ID","Behavior","Family"),
                      measure.vars = c("biPC1", "biPC2", "biPC3", "biPC4"))

melt.beh$Family <- factor(melt.beh$Family , levels=c("Nymphalidae", "Hesperiidae", "Lycaenidae", "Pieridae", "Papilionidae"))
write.csv(melt.beh, "beh_for_JMP_May15.csv")


beh.tot <- glm(Behavior ~ biPC1 + biPC2 + biPC3 + biPC4 + Site, data=bi.combo, family=binomial)
summary(beh.tot)
anova(beh.tot, test = "Chisq")


beh.hes<-filter(bi.combo,Family=="Hesperiidae")
beh.pap<-filter(bi.combo,Family=="Papilionidae")
beh.nym<-filter(bi.combo,Family=="Nymphalidae")
beh.pie<-filter(bi.combo,Family=="Pieridae")
beh.lyc<-filter(bi.combo,Family=="Lycaenidae")
#Hesperiidae
beh.glm.hes <- glm(Behavior ~ biPC1 + biPC2 + biPC3 + biPC4 + Site, data=beh.hes, family=binomial)
summary(beh.glm.hes)
anova(beh.glm.hes, test = "Chisq")


#Nymphalidae
beh.glm.nym <- glm(Behavior ~ biPC1 + biPC2 + biPC3 + biPC4 + Site, data=beh.nym, family=binomial)
summary(beh.glm.nym)
anova(beh.glm.nym, test = "Chisq")


#Lycaenidae
beh.glm.lyc <- glm(Behavior ~ biPC1 + biPC2 + biPC3 + biPC4 + Site, data=beh.lyc, family=binomial)
summary(beh.glm.lyc)
anova(beh.glm.lyc, test = "Chisq")


###Not enough data
#Pieridae
beh.glm.pie <- glm(Behavior ~ biPC1 + biPC2 + biPC3 + biPC4 + Site, data=beh.pie, family=binomial)
summary(beh.glm.pie)
anova(beh.glm.pie, test = "Chisq")

#Papilionidae
beh.glm.pap <- glm(Behavior ~ biPC1 + biPC2 + biPC3 + biPC4 + Site, data=beh.pap, family=binomial)
summary(beh.glm.pap)
anova(beh.glm.pap, test = "Chisq")
```